---
title: "Intro to Modeling"
format: html
editor: visual
---

```{r}
library(readr)
library(dplyr)
library(stringr)
library(lubridate)
library(tidyr)
library(ggplot2)
library(ggcorrplot)
library(tidymodels)

bike_raw <-  read_csv(
  "./Data/SeoulBikeData.csv",
  col_names = TRUE,
  locale = locale(encoding = 'latin1')
)
```

## EDA

```{r}
bike_raw |> 
  summarise(across(everything(), ~sum(is.na(.))))
```

We observe no missing values.

```{r}
str(bike_raw)
```

We see the types generally make sense, so we'll only convert the date column (and create factors from the string variables) but rename all with a short function.

```{r}
format_column_name <- function(name) {
  name |> 
    str_remove_all("[^[:alnum:] ]")  |>
    str_squish() |> 
    str_to_lower() |> 
    str_replace_all(" ", "_") |> 
    str_replace_all("temperaturec", "temperature_c")
}

bike_tbl <-  bike_raw |> 
  rename_with(format_column_name) |> 
  mutate(date = dmy(date)) |> 
  mutate(across(where(is.character), as.factor))

bike_tbl |>
  summarize(across(where(is.numeric), list("mean" = mean, "sd" = sd), .names = "{.fn}_{.col}"))

bike_tbl |>
  group_by(seasons, holiday, functioning_day) |> 
  summarize(across(where(is.numeric), list("mean" = mean, "sd" = sd), .names = "{.fn}_{.col}"))

table(bike_tbl$seasons)
table(bike_tbl$holiday)
table(bike_tbl$functioning_day)
```

We can get a sense of trends by group, in particular that non-functioning days will have 0 rentals.

```{r}
bike_tbl <- bike_tbl |> 
  filter(functioning_day == "Yes") |> 
  select(-functioning_day)
```

Now we want to only have one observation per day and use the daily average or sum of certain statistics.

```{r}
bike_tbl <- bike_tbl |> 
  group_by(date, seasons, holiday) |> 
  mutate(across(c(rented_bike_count, rainfallmm, snowfall_cm), list("sum" = sum), .names = "{.col}")) |> 
  mutate(across(c(temperature_c, humidity, wind_speed_ms, visibility_10m, dew_point_temperature_c, solar_radiation_mjm2), list("mean" = mean), .names = "{.col}")) |> 
  filter(row_number()==1) |> 
  select(-hour)
```

Now we summarize again, with a few plots for good measure.

```{r}
bike_tbl |>
  ungroup() |> 
  summarize(across(where(is.numeric), list("mean" = mean, "sd" = sd), .names = "{.fn}_{.col}"))

bike_tbl |>
  group_by(seasons, holiday) |> 
  summarize(across(where(is.numeric), list("mean" = mean, "sd" = sd), .names = "{.fn}_{.col}"))

bike_tbl |> 
  ungroup() |> 
  select(where(is.numeric)) |> 
  cor() |>
  ggcorrplot(hc.order = TRUE, type = "lower", lab = TRUE)

bike_tbl |> 
  ggplot(aes(x = rented_bike_count)) +
  geom_histogram()

bike_tbl |> 
  ggplot(aes(x = rented_bike_count, fill = holiday)) +
  facet_wrap("seasons") +
  geom_histogram()

bike_tbl |> 
  ggplot() +
  geom_point(aes(x = temperature_c, y = rented_bike_count))

bike_tbl |> 
  ggplot() +
  geom_line(aes(x = date, y = rented_bike_count))
```

We see a few trends:

-   Weather appears to impact rentals (with more summery weather correlated positively with rentals), but could also be correlated with season.

-   Winter almost creates a "secondary mode" in the rental histogram since it has so many low-rental days.

-   Holidays don't seem to increase bike rentals, but also don't obviously decrease them.

## Modeling

First we'll split up our data as desired.

```{r}
set.seed(123456)
bike_split <- initial_split(bike_tbl, strata = seasons)

bike_cv_folds <- vfold_cv(training(bike_split), 10)
```

Now we build recipes.

```{r}
recipe_1 <- recipe(rented_bike_count ~ ., data = training(bike_split)) |> 
  step_mutate(day_type = factor(ifelse(wday(date, label = TRUE) %in% c("Sat", "Sun"), "Weekend", "Weekday"))) |> 
  step_normalize(where(is.numeric)) |> 
  step_dummy(c("seasons", "holiday", "day_type")) |> 
  step_rm(date)

wf_1 <- workflow() |> 
  add_recipe(recipe_1) |> 
  add_model(linear_reg() |> set_engine("lm"))

wf_1 |> 
  fit_resamples(bike_cv_folds) |> 
  collect_metrics()

recipe_2 <- recipe(rented_bike_count ~ ., data = training(bike_split)) |> 
  step_mutate(day_type = factor(ifelse(wday(date, label = TRUE) %in% c("Sat", "Sun"), "Weekend", "Weekday"))) |> 
  step_normalize(where(is.numeric)) |> 
  step_dummy(seasons, holiday, day_type) |> 
  step_interact(terms = ~ starts_with("holiday"):starts_with("seasons")) |> 
  step_interact(terms = ~ starts_with("seasons"):temperature_c) |> 
  step_interact(terms = ~ temperature_c:rainfallmm) |> 
  step_rm(date)

wf_2 <- workflow() |> 
  add_recipe(recipe_2) |> 
  add_model(linear_reg() |> set_engine("lm"))

wf_2 |> 
  fit_resamples(bike_cv_folds) |> 
  collect_metrics()

recipe_3 <- recipe(rented_bike_count ~ ., data = training(bike_split)) |> 
  step_mutate(day_type = factor(ifelse(wday(date, label = TRUE) %in% c("Sat", "Sun"), "Weekend", "Weekday"))) |> 
  step_mutate(across(where(is.numeric), ~ .x^2, .names = "{.col}_quad")) |> 
  step_normalize(where(is.numeric) & !ends_with("_quad")) |> 
  step_dummy(seasons, holiday, day_type) |> 
  step_interact(terms = ~ starts_with("holiday"):starts_with("seasons")) |> 
  step_interact(terms = ~ starts_with("seasons"):temperature_c) |> 
  step_interact(terms = ~ temperature_c:rainfallmm) |> 
  step_rm(date)

wf_3 <- workflow() |> 
  add_recipe(recipe_3) |> 
  add_model(linear_reg() |> set_engine("lm"))

wf_3 |> 
  fit_resamples(bike_cv_folds) |> 
  collect_metrics()
```

It looks like our error is minimized in model 3, so we'll use that for the last part.

```{r}
test_run <- wf_3 |> 
  last_fit(bike_split)

test_run |> 
  collect_metrics()
test_run |> 
  extract_fit_parsnip() |> 
  tidy()
```

It looks like that model does pretty well here. Our coefficients are interesting too, to see which variables are used heavily in the model.
